A Hitchhiker's Guide to Reproducible Research in R

### 2021-07-29 manuscript.Rmd
start incorporating Andreas' draft discussion

---
title: A Hitchhiker's Guide to Reproducible Research in R
author:
  - name: Aaron Peikert
    affil: 1, *
    orcid: 0000-0001-7813-818X
  - name: Caspar J. Van Lissa
    affil: 2, 3
    orcid: 0000-0002-0808-5024
  - name: Andreas M. Brandmaier
    affil: 1, 4
    orcid: 0000-0001-8765-6982
affiliation:
  - num: 1
    address: |
      Center for Lifespan Psychology---Max Planck Institute for Human Development
      Lentzeallee 94, 14195 Berlin, Germany
  - num: 2
    address: |
      Department of Methodology & Statistics---Utrecht University faculty of Social and Behavioral Sciences, Utrecht, Netherlands
  - num: 3
    address: |
      Open Science Community Utrecht, Utrecht, Netherlands
  - num: 4
    address: |
      Max Planck UCL Centre for Computational Psychiatry and Ageing Research
      Berlin, Germany and London, UK
# firstnote to eighthnote
correspondence: |
  peikert@mpib-berlin.mpg.de
journal: psych
type: tutorial
status: submit
bibliography: temp.bib
simplesummary: |
  Reproducibility has long been considered integral to the scientific method.
  Something is called reproducible when independend researcher arrive at the same results from the same data.
  Until recently, detailed descriptions of methods and analyses were the primary instrument for ensuring scientific reproducibility.
  Technological advancements now enable scientists to achieve a much higher standard;
  one in which any individual can be granted access to a digital research repository, and reproduce the analyses from raw data to final manuscript with a single command.
  This method has far-reaching implications for scientific archiving,
  reproducibility and replication,
  scientific productivity,
  and the credibility and reliability of scientific findings.
  One obstacle preventing the widespread adoption of this method is that the underlying technological advancements
  are complicated to use.
  This paper introduces `repro`; an R-package that guides researchers in the installation and use of the tools required for computational reproducibility.
  It includes a tutorial on how to make a research project reproducible with the aid of `repro`.
  Finally, it introduces novel applications of reproducibility tools for scientific research;
  including the preregistration of study plans as reproducible computer code.
  This circumvents the shortcomings of ambiguous preregistrations that allow for researcher degrees of freedom,
  because computer code describes study plans more precisely than prose.
  Reproducibility made convenient with automation has hence a wide range of applications to accelerate scientific progress.
abstract: "`r tryCatch(trimws(readr::read_file(here::here('abstract.Rmd'))))`"
keywords: |
  keyword 1; keyword 2; keyword 3 (list three to ten pertinent keywords specific 
  to the article, yet reasonably common within the subject discipline.).
acknowledgement: |
  The authors received no financial support for the research, authorship, and/or publication of this article.
authorcontributions: |
  Aaron Peikert took the lead in writing and provided the initial draft of the manuscript.
  Andreas Brandmaier and Caspar J. Van Lissa contributed further ideas, critical feedback, and revisions of the original manuscript.
  Furthermore, we would like to thank Maximilian Stefan Ernst (not an author) for his contributions to the code for the simulation study.
conflictsofinterest: |
  The authors declare no conflict of interest.
abbreviations:
  - short: PAC
    long: Preregistration as 
  - short: Gb
    long: Gigabyte
  - short: Kb
    long: Kilobyte
  - short: CRAN
    long: Comprehensive R Archive Network
repro:
  packages:
    - tidyverse
    - usethis
    - gert
    - aaronpeikert/repro@c3cae1f
    - here
    - rstudio/webshot2@f62e743
    - targets
    - renv
    - slider
    - patchwork
    - knitr
    - pander
    - lavaan
    - furrr
    - future.batchtools
    - rticles
    - moments
    - report
  scripts:
    - R/simulation.R
    - R/simulation_funs.R
    - R/link.R
  data:
    - data/simulation_results.csv
output:
  bookdown::markdown_document2:
    base_format: rticles::mdpi_article
header-includes:
   - \usepackage[labelformat=empty]{caption}
   - \DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}, numbers=left}
---

@@

One impediment to the widespread adoption of a standard for reproducible research is that learning to use the required tools can be time-intensive.
To lower the threshold to adoption, the R-package `repro` introduces helper functions that wrap more complicated (and powerful) tools.
These so-called wrappers guide end-users in the use of reproducibility tools, provide feedback about what the computer does, and suggest what the user should do in layman's terms.
We hope this makes reproducibility tools more accessible by enabling beginner-level users to detect their system's state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
These wrappers are merely an assistance system, and as users get more and more comfortable, they can use the underlying tools directly to solve more complex problems.

This tutorial assumes that the user is working predominantly in R, with the help of RStudio.
It describes basic steps that we expect to apply to all users. Your specific situation might involve additional, more specialized steps. After completing the tutorial, you should be able to customize your workflow accordingly.
The first step is to install the required software.
Start R Studio<!--CJ I am concerned that new users may not install R if you only mention RStudio.--> and install the package [`repro`](https://github.com/aaronpeikert)[@R-repro]. 
It will assist you while you follow the tutorial.

```{r, eval=FALSE}
# repro is not on CRAN yet
options(
  repos = c(aaronpeikert = 'https://aaronpeikert.r-universe.dev',
            CRAN = 'https://cloud.r-project.org')
)
install.packages('repro')
```

An example of the assistance that `repro` provides are the check functions.
These verify that you indeed have installed and setup the required pieces of software for this workflow (here, we use the double-colon operator to access functions from the `repro` namespace; we chose to do so to make clear which functions are native `repro` functions) :

@@

These functions check whether specific dependencies are available on the users' system, and if not, explain what further action is needed to obtain it.
Sometimes they ask the user to do something, for example, the following happens if you were a Windows user who does not have Git installed:

@@

The messages from `repro` try to take the user by the hand to help them solve problems.
These messages are adjusted to your specific operating system and installed dependencies.
Before you continue, we ask you to run the above commands to check Git, Make, and Docker - both to become familiar with the functionality of the `check_*()` functions, and to prepare your system for creating reproducible projects.
If necessary, follow any instructions presented until all checks pass.

## Implementing version control

Now that your system is set up for this tutorial, we will introduce you to version control with Git.
Git tracks files within one project folder.
We start by creating such a project folder with RStudio by clicking the menu item:

@@

Git tracks any files that you manually add to the repository.
To make sure you do not accidentally add files that you do not wish to share (e.g., privacy sensitive data),
you can list specific files that you do not want to track in the `.gitignore` file.
You can add something to the `.gitignore` file directly or with this command:

```{r, eval=FALSE}
usethis::use_git_ignore("private.md")
```
Now the file `private.md` will not be committed to Git and will not be made public.

Especially for new users, it is recommended to follow this procedure and exclude sensitive files before proceeding.
When you are ready, you can begin tracking your files using Git by running:

@@

For Git to recognize changes to a given file, you have to stage and then commit these changes (this is the basic save action for a project snapshot).
One way to do this is through the visual user interface in the RStudio Git pane.
Click on the empty box next to the file you want to stage. A check mark appears to indicate that the file is staged.
After you have staged all files you want, click on the commit button, explain in the commit message why you made those changes, and then click on commit.
This stores a snapshot of the current state of the project.

The files you created and the changes you made still have not left your computer.
All snapshots are stored in a local repository in your project folder.
To back up and/or share the files online, you can push your local repository to a remote repository.
While you can choose any Git service (like GitLab or BitBucket), we will use GitHub in this tutorial.
Before you upload your project to Git, you need to decide if you would like the project publically accessable (viewable by anyone, editable by selected collaborators) or if you want to keep it private (only viewable and editable by selected collaborators).
To upload the project publically to GitHub use:

@@

To upload it privately:

@@

## Dynamic document generation

Now that you have created a version controlled project, we will proceed with dynamic document generation.
A dynamic document has three elements:

1. Text in natural language (prose; e.g., a scientific paper or presentation)
2. Executable code (e.g., analyses)
3. Metadata (e.g. title, authors, document format)

R Markdown is a type of dynamic document well-suited to the RStudio user interface.
The natural language text in an R Markdown file can be formatted using Markdown (see @xieMarkdownDefinitiveGuide2019 for technical details and @xieMarkdownCookbook2020 for practical guidance).
The code mostly consists of R-code (although other programming languages are supported, like Python, C++, Fortran, etc).
The following example serves to illustrate the Markdown syntax. 
It shows how to create a heading, a word in bold font, and a list of several items in Markdown:

@@

One advantage of this type of markup for formatting is that it can be rendered to many different output formats - both in terms of file types, like `.docx`, `.html`, `.pdf`, and in terms of style, e.g. specific journal requirements.
Another advantage is that, because R Markdown files are in plain text, they are more suitable for version control using Git.
Some users might find it easier to activate the "Visual Editor" of R Studio (Ctrl + Shift + F4 or click on the icon that resembles drawing materials or a compass <!--CJ Add picture-->in the upper right corner of the R Markdown document) which features more graphical elements like a traditional word processor but still creates an R Markdown underneath with all of its flexibility.
The visual editor has some additional benefits, such as promoting best practices (for example, each sentence should be written on a newline, which eases tracking changes across versions)
and easing the generation of citations and references to Tables and Figures.

Now that you are familiar with the basics of Markdown formatting, we turn our attention to including code into the text.
Code is separated by three backticks (and the programming language in curly braces) like here:

@@

`r ''````{r}
# this is R code
1 + 1
```
````
The hotkey [Control]+[Alt]+[i] inserts a block of code in the file.
The results of code enclosed in such backticks will be dynamically inserted into the document (depending on specific settings).
This means that, whenever you render the R Markdown to its intended output format, the code will be executed and the results updated.
The resulting output document will be static, e.g. a pdf document, and can be shared wherever you like, e.g on a preprint server.

Once the R Markdown file has been rendered to a static document, the resulting file is decoupled from the R Markdown and the code that created it.
This introduces a risk that multiple versions of the static document are disseminated, each with slightly different results.
To avoid ambiguity, we therefore recommend referencing the identifier of the Git commit at the time of rendering in the static document.
Simply put, a static document should link to the version of the code that was used to create it.
The `repro` package comes with the function `repro::current_hash()` for this purpose.
This document was created from the commit with the hash ``r repro::current_hash()``.

Now that you know how to write text and R code in an R Markdown, you need to know about metadata (also called: YAML front matter).
These metadate contain information about the document, like the title and the output format.
Metadata are placed at the beginning of the document and are separated by three dashes from the document body; the following example is a full markdown document where the header is in lines 1--6.

````markdown
---
title: A Hitchhiker's Guide to Reproducible Research in R
author: Aaron Peikert, Caspar J. van Lissa and Andreas M. Brandmaier
abstract: A tutorial about doing the same thing more than once.
output: html_document
---

@@

Some metadata fields are self-explanatory (like the author field), and exist across all output formats (like the title field).
However, R Markdown has been extended to hundreds of document formats which are often highly configurable by their metadata.
Some R packages add additional metadata fields.
The `repro` package adds fields to the metadata to help make the research project reproducible.
These fields are used to list all dependencies of the research project.
This includes R scripts, data files, and external packages.
The format is as follows (see everything below the line *repro:*):

@@

This information makes explicitly declares the project's dependencies on files and R packages.
`repro` uses this information to infer a Makefile for the dependencies on other files and a Dockerfile that includes all required packages.
These two files together form the basis for consistency within a research project and consistency across different systems.<!--CJ Consider explaining this in the theoretical intro-->
At this point, we would like to emphasize that a preqrequisite for reproducibility is that every step taken in an analysis can be automatically reproduced.<!--CJ This comes way to late. Shouldbe in the opening paragraph or something.-->
This means there must be no manual processing such as copy and pasting, manual modification of data files (e.g., manually removing observations or manually recoding variables). 
Ideally, all data preprocessing is well-documented and gathered in a separate R file, which is then called before a Markdown report is generated (that is, the preprocessing script is added in the metadata header under _scripts:_).<!--CJ These points describe the formal definition of reproducibility used in this paper, and the goal of the workflow. That needs to be explained in the introduction. If you only find out what you are working toward halfway through the tutorial - that's too late!-->

The function `repro::automate()` converts the metadata from all R Markdown files in the project (all files with the ending `.Rmd`) to a Makefile and a Dockerfile.
Together, these files allow users (including your future self) to reproduce every step in the analysis automatically.
It is important to re-run `repro::automate()` whenever you change the `repro`-metadata, change the output format, or add a new R Markdown file to the project, to keep the Makefile and Dockerfile up to date.
The function `repro::automate()` does not overwrite manual changes<!--CJ What does that mean?-->, and there is no harm in running it too often.<!-- CJ Can that not be automated? Or at least provide a suggestion when to run it?-->
Other than the Makefile and the Dockerfile, which are created in the document root path, repro generates a few more files in the `.repro` directory (which we will explain in detail later), all of which you should add and commit to Git. <!--AB: why is this addition not automated? Feels strange given that this is our recommended best practice?! CJ: Agreed.-->

<!-- CJ What i'm missing here is something to wrap up the tutorial. Do users now have a project that could be reproduced? The tutorial should not end until that is the case.-->

## Reproducing a project

<!--CJ Suggestion: Instead of explaining reproduction right in the middle of the tutorial for setting up a project, add a later section called "reproducing a project", and provide an example project for users to reproduce.-->
When you or someone else who received your project want to reproduce it, they can ask repro to explain how this should be done:

@@

Importantly, a reproducible research project created with `repro` does not have the `repro` package itself as a dependency.<!--CJ This belongs in the intro, not the tutorial. It is part of explaining the benefits of repro.-->
These projects will remain reproducible irrespective of whether `repro` will be maintained and remain accessible in future.
Users do not have to have repro installed to reproduce a project, in fact, they do not even need to have R installed because the entire project can be rebuilt inside of a container that has R installed.
The only required software for reproducing a project is Docker, assuming users know how to build a Docker image and run Make within the container.
However, if they have installed Make additionally to Docker, they do not need to know how to use Docker and can simply rely on the two Make commands `make docker` and `make -B DOCKER=TRUE`.
The commands can be pasted into a terminal and will reproduce the project in many cases.

@@

This section is only for advanced users who want to overcome some limitations of `repro`.
If you read this paper the first time you probably want to skip this section.
As explained above, `repro` is merely a simplified interface that enables a high level of reproducibility for simple projects.
The simplification of Docker and Make are possible because of two restrictions.
Users who ask themselves either "How can I install software dependencies outside of R in the Docker image?" or "How can I express complex dependencies between files, e.g. hundreds of data files are preprocessed and combined?") need to be aware of these restriction and require a deeper understanding of the inner workings of `repro`.
Other users may safely skip this section, or return to it if they encounter such challenges.

@@

We believe that pairing PAC with the here presented workflow offers five advantages over classical preregistration.
First, a PAC removes any ambiguity regarding the translation of an verbal analysis plan into code.
Second, while being rigorous, it offers the flexibility to incorporate data-dependent decisions if they can be formulated as code.
Researchers can, for example, formulate conditions under which they prefer one analysis type over the other.
For example, if distributional assumptions are not met, the code may branch out to employ robust methods; or, an analysis may perform automated variable selection mechanisms before running the final model.
Another example of data dependent decisions are more explorative analyses, i.e. explorative factor analysis or machine learning.
With PAC researchers can precisely calibrate how much they want to explore.
When communicating a finding as explorative, it may still make a difference if tens or hundreds of variables were considered.
Third, a PAC is more comprehensive by design because its completeness can be empirically checked with simulated data.
We believe it is easier with PAC to detect missing steps or ambiguous decisions.
Because of its strict structure, PAC can therefore more effectively minimize researchers degrees of freedom than a standard preregistration [@bakkerEnsuringQualitySpecificity2020].
Fourth, deviations from the preregistration are visible because they are reflected in changes to the code, which are managed and tracked with version control.
Fifth, the preregistration is merely a development stage of the final manuscript, thus sparing authors from writing and editors and reviewers from evaluating two separate documents.

@@

<!--AB: TODO add a quick summary -->

@@

PACs come with a challenge though.
In contrast to a reproducible script, which should work on the same data, a PAC must work on different data.
We call the property of code that can run on different inputs from a similar context and produce valid outputs "reusability".
This property is based on reproducibility, but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] such that it is _build-for-reuse_ [@al-badareenReusableSoftwareComponents2010].
The reproducible workflow we present here is heavily automated and hence promotes reusability.
Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019] and therefore this workflow fits well together with other innovations that employ automation, like machine readable hypothesis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
Automated and reusable research projects promises a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].<!--AB very nice!!-->

@@

<!--AB: add limitations; move the t-test discussion on non-optimality here TODO-->
Please note that the procedure of running different test procedures based on a fixed skewness threshold was created for didactic purposes and it is not necessarily optimal for adressing your research question.
The t-test and Mann–Whitney–Wilcoxon test are arguably the most often used hypothesis test [@fagerlandTtestsNonparametricTests2012; @hortonStatisticalMethodsJournal2005 reports that 26% of all studies employ a t-test and 27% employ a rank based alternative in the New England Journal of Medicine in 2005].
The analytical strategy presented here is in fact suboptimal with regard to several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]). <!--AB can we move this to a different place? Maybe this would actually better fit to the discussion/limitation section!-->

@@

<!--AB: two lines of what the future of the package, of such standards, and prereigstration is. Maybe say something that preregistration is not always possible or wanted but that reproducibility is almost always a good idea?-->

>>>>

---
title: A Hitchhiker's Guide to Reproducible Research in R
author:
  - name: Aaron Peikert
    affil: 1, *
    orcid: 0000-0001-7813-818X
  - name: Caspar J. Van Lissa
    affil: 2, 3
    orcid: 0000-0002-0808-5024
  - name: Andreas M. Brandmaier
    affil: 1, 4
    orcid: 0000-0001-8765-6982
affiliation:
  - num: 1
    address: |
      Center for Lifespan Psychology---Max Planck Institute for Human Development
      Lentzeallee 94, 14195 Berlin, Germany
  - num: 2
    address: |
      Department of Methodology & Statistics---Utrecht University faculty of Social and Behavioral Sciences, Utrecht, Netherlands
  - num: 3
    address: |
      Open Science Community Utrecht, Utrecht, Netherlands
  - num: 4
    address: |
      Max Planck UCL Centre for Computational Psychiatry and Ageing Research
      Berlin, Germany and London, UK
# firstnote to eighthnote
correspondence: |
  peikert@mpib-berlin.mpg.de
journal: psych
type: tutorial
status: submit
bibliography: temp.bib
simplesummary: |
  A Simple summary goes here.
abstract: "`r tryCatch(trimws(readr::read_file(here::here('abstract.Rmd'))))`"
keywords: |
  keyword 1; keyword 2; keyword 3 (list three to ten pertinent keywords specific 
  to the article, yet reasonably common within the subject discipline.).
acknowledgement: |
  The authors received no financial support for the research, authorship, and/or publication of this article.
authorcontributions: |
  Aaron Peikert took the lead in writing and provided the initial draft of the manuscript.
  Andreas Brandmaier and Caspar J. Van Lissa contributed further ideas, critical feedback, and revisions of the original manuscript.
  Furthermore, we would like to thank Maximilian Stefan Ernst (not an author) for his contributions to the code for the simulation study.
conflictsofinterest: |
  The authors declare no conflict of interest.
abbreviations:
  - short: PAC
    long: Preregistration as 
  - short: Gb
    long: Gigabyte
  - short: Kb
    long: Kilobyte
  - short: CRAN
    long: Comprehensive R Archive Network
repro:
  packages:
    - tidyverse
    - usethis
    - gert
    - aaronpeikert/repro@c3cae1f
    - here
    - rstudio/webshot2@f62e743
    - targets
    - renv
    - slider
    - patchwork
    - knitr
    - pander
    - lavaan
    - furrr
    - future.batchtools
    - rticles
    - moments
    - report
  scripts:
    - R/simulation.R
    - R/simulation_funs.R
    - R/link.R
  data:
    - data/simulation_results.csv
output:
  bookdown::markdown_document2:
    base_format: rticles::mdpi_article
header-includes:
   - \usepackage[labelformat=empty]{caption}
   - \DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}, numbers=left}
---

@@

The above mentioned tools require much expertise that is beyond the traditional curriculum of researchers.
To aid researcher adopting these tools we rely on or have developed helpers that wrap more complicated (and powerful) tools.
These wrappers use standard tools but provide feedback for the user about what the computer does and what the user should do in layman's terms and are accessible from within R.
We hope this makes reproducibility tools more accessible by enabling beginner-level users to detect their systems state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
These alternative tools are merely an assistance system, and as users get more and more comfortable, they can use the underlying tools directly to solve more complex problems.

Before we can start creating reproducible research projects, we need a computer with a working R Studio installation. 
Start R Studio and install the package [`repro`](https://github.com/aaronpeikert)[@R-repro] so that it can act as an assistance systems while you follow the tutorial:

```{r, eval=FALSE}
# repro is not on CRAN yet
install.packages('repro', repos = 'https://aaronpeikert.r-universe.dev')
```

An example of the assistance that `repro` provides are the check functions.
They verify that you indeed have installed and setup the required pieces of software for this workflow (here, we use the double-colon operator to access functions from the `repro` namespace; we chose to do so to make clear which functions are native `repro` functions) :

@@

They detect the users' system state and explain if further action is needed.
Sometimes they ask the user to do something, for example, the following happens if you were a Windows user who does not have Git installed:

@@

The messages emitted by `repro` and other packages we present try to take the user by the hand to help them solve problems.
This tutorial describes steps that we expect to apply to all users; the messages from the packages are adjusted to your specific situation (that is, operating system and the specific problem of missing software) and complement the tutorial in this way.
Before you continue, we ask you to run the above commands to check Git, Make, and Docker.
If necessary, follow the instructions of repro to install them until the check functions are all fine.

The first concept we want to introduce is version control with Git.
Git tracks files within one project folder.
We start by creating such a project folder with RStudio by clicking the menu item:

@@

You can prevent Git from accidentally committing files that you do not want to track with the `.gitignore` file.
You can add something to the `.gitignore` file directly or with this command:

```{r, eval=FALSE}
usethis::use_git_ignore("private.md")
```

Now the file `private.md` will not be committed to Git and will not be made public.

After you have excluded sensitive files (like not yet anonymized data files), you can activate Git with:

@@

For Git to recognize changes to a given file, you have to stage and then commit these changes (this is the basic save action for a project snapshot).
The easiest way to do this is to use the RStudio Git pane.
Click on the empty box next to the file you want to stage, and when a tick appears, which means that the file is staged.
After you have staged all files you want, click on the commit button, explain in the commit message why you made those changes, and then click on commit.

The files you created and the changes you made still have not left your computer.
All snapshots are stored in a local repository in your project folder.
To back up and share the files, you can publicly upload those changes to a public repository.
To synchronize your local repository with a remote repository from the GitHub service, issue the following command:

@@

If you are not yet ready to share your project publicly but you want to synchronize your local repository to a remote repository (for example, as a simple backup service), use:

@@

Now that you have created a project that is under version control, we continue with adding dynamic document generation.
Adynamic document has three elements:

1. Text in a natural language
2. Executable code 
3. Metadata

The text of an R Markdown is written in Markdown (see @xieMarkdownDefinitiveGuide2019 for technical details and @xieMarkdownCookbook2020 for practical guidance).
The following example serves to illustrate the Markdown syntax. 
It shows how to create a heading, a word in bold font, and a list of several items in Markdown:

@@

This type of markup for formatting, instead of the more visual approach of e.g. Microsoft Word, is compatible with many output formats and is easier to put under version control.
Some users might find it easier to activate the "Visual Editor" of R Studio (Ctrl + Shift + F4 or click on the icon that resembles drawing materials or a compass in the upper right corner of the R Markdown document) which features more graphical elements like a traditional word processor but still creates an R Markdown underneath with all of its flexibility.
The visual editor has the additional benefit of promoting best practices (for example, each sentence should be written on a newline, which eases tracking changes across versions).
Among other features, it eases the generation of citations and cross references.

Now that you are familiar with the basics of writing text in Markdown, we turn our attention to including code into the text.
Code is separated by three backticks (and the programming language in curly braces) like here:

@@

`r ''````{r}
# this is R code
1 + 1
```
````

The results of your code are than dynamically inserted into the document.
Dynamically means that whenever you turn the R Markdown into its output format the code is executed again and if you changed something the results are updated.
The output document is than static, e.g. a pdf document, and can be shared wherever you like, e.g on a preprint server.
This static document is somewhat decoupled from the R Markdown and the rest of the code that did create it.
It is quite possible that several slightly different versions are therefore circulating.
We therefore recomend to connect the static document to the version control system by dynamically including the identifier of the identifier of the Git commit, called hash, in the document.
The `repro` package comes with the function `repro::current_hash()` for that purpose.
This document was created from the commit with the hash ``r repro::current_hash()``.

Now that you know how to write text and R code in an R Markdown, you need to know about metadata, like the title or the output format. Metadata are placed at the beginning of the document and are separated by three dashes from the document body; the following example is a full markdown document where the header is in lines 1--6.

````markdown
---
title: A Hitchhiker's Guide to Reproducible Research in R
author: Aaron Peikert, Caspar J. Van Lissa and Andreas M. Brandmaier
abstract: A tutorial about doing the same thing more than once.
output: html_document
---

@@

Most metadata fields are self-explanatory (like the author field). 
Sometimes, R package may extend the metadata fields for their own purposes.
Certain output formats may require specific  metadata. 
For example, the `rticles`-package that offers templates for authoring scientific manuscripts comes with extra fields that are described in the package documentation and offer extra options for flexible formatting of documents.
R Markdown has been extended to hundreds of document formats which are often highly configurable by their metadata.
The `repro` package uses the metadata for hints to make the research project reproducible.
Researchers should list all dependent resources of their research project in the metadata, that is, R scripts, data files, and external packages in the following format (see everything below the line *repro:*):

@@

This information makes internal dependencies to other files in the project and external dependencies to R packages explicit.
`repro` can use this information to infer a Makefile for the dependencies on other files and a Dockerfile that includes all required packages.
These two files together form the basis for consistency within a research project and consistency across different systems.
At this point, we would like to emphasize that a preqrequisite for reproducibility is that every step taken in an analysis can be automatically reproduced.
This means there must be no manual processing such as copy and pasting, manual modification of data files (e.g., manually removing observations or manually recoding variables). 
Ideally, all data preprocessing is well-documented and gathered in a separate R file, which is then called before a Markdown report is generated (that is, the preprocessing script is added in the metadata header under _scripts:_).
In everyday work, the simplest solution is to now call the function `repro::automate()` which both creates a Makefile and a Dockerfile from all R Markdown files in the project (all files with the ending `.Rmd`) based on the respective repro metadata entries.
Every time you change the `repro`-metadata, change the output format, or add a new R Markdown file to the project, the Makefile and Dockerfile are potentially outdated, and you should run `repro::automate()` to keep them up to date.
The function `repro::automate()` does not overwrite manual changes, and there is no harm in running it too often.
Other than the Makefile and the Dockerfile, which are created in the document root path, repro generates a few more files in the `.repro` directory (which we will explain in detail later), all of which you should add and commit to Git. <!--AB: why is this addition not automated? Feels strange given that this is our recommended best practice?!-->

When you or someone else who received your project want to reproduce it, they can ask repro to explain how this should be done:

@@

Importantly, a reproducible research project created with `repro` does not have the `repro` package itself as a dependency.
These projects will remain reproducible irrespective of whether `repro` will be maintained and remain accessible in future.
Users do not have to have repro installed to reproduce a project, in fact, they do not even need to have R installed because the entire project can be rebuilt inside of a container that has R installed. 
The two Make commands `make docker` and `make -B DOCKER=TRUE` together with an installation of Make and Docker are sufficient.
The commands can be pasted into a terminal and will reproduce the project in many cases.
<!--AB: I said this before. I think it really should be possible to execute this command from within R. Never underestimate the troubles people can have with the terminal. I think we discussed this before. Please add a function repro::rerun(execute=TRUE) or -- if we rename it -- repro::reproduce(execute=TRUE). -->

@@

Some projects, however, are not as easy to make reproducible.
As explained above, `repro` is merely a simplified interface that guarantees a high level of reproducibility for simple projects.
These simplifications are possible because of two restrictions.
Users who ask themselves either "How can I install software that is not an R package in the Docker image?" or "How can I express complex dependencies between files, e.g. raw data turned into preprocessed data required by analysis?" need to be aware of these restriction and require a deeper understanding of the inner workings of `repro`.
All other users may safely skip this section and may return to it if they become limited by restriction.

@@

We believe that pairing PAC with the here presented workflow offers five advantages over classical preregistration.
First, a PAC removes any ambiguity regarding the translation of an verbal analysis plan into code.
Second, while being rigorous, it offers the flexibility to incorporate data-dependent decisions if they can be formulated as code.
Researchers can, for example, formulate conditions under which they prefer one analysis type over the other.
For example, if distributional assumptions are not met, the code may branch out to employ robust methods; or, an analysis may perform automated variable selection mechanisms before running the final model.
Third, a PAC is more comprehensive by design because its completeness can be empirically checked with simulated data.
We believe it is easier with PAC to detect missing steps or ambiguous decisions.
Because of its strict structure, PAC can therefore more effectively minimize researchers degrees of freedom than a standard preregistration [@bakkerEnsuringQualitySpecificity2020].
Fourth, deviations from the preregistration are visible because they are reflected in changes to the code, which are managed and tracked with version control.
Fifth, the preregistration is merely a development stage of the final manuscript, thus sparing authors from writing and editors and reviewers from evaluating two separate documents.

@@

## Summary

We demonstrated how the R package `repro` supports researchers in creating reproducible repositories and reproducible manuscripts.
These are important building blocks for transparent and cumulative science because they enable others to reproduce statistical and computational results and reports later in time and on different computers.
The workflow we present here rests on four software solutions, version control, dynamic document generation, dependency tracking and software management to all but guarantee reproducibility.

We then proceeded to show preregistration as code (PAC) as one example of how such rigorous and automated reproducibility workflow may enable other innovations.
In a PAC researcher write all analysis code and a reproducible manuscript that already renders the entire reporting of statistical and computational results before they gather data.
These results are based on simulated data, such that there is no danger of peeking at the real data while writing the scripts.
This way, all researcher's degrees of freedom are eliminated.
Once real data is gathered, the reproducible manuscript is (re-)created with the real data.
Because the same dynamic document serves as preregistration and final manuscript, researcher, reviewers and editors only have to revise one document.
To editors, reviewers and readers it is obvious how the researchers deviated from the preregistration because everything is under version control.
The likelihood of those deviations, while unavoidable, is decreased by using simulated data to verify the suitability of the analysis plan.
This analysis plan may contain data depended decisions to further decrease the likelihood that researcher are unable stick to their preregistration

## Limitations of repro

`repro`'s primary goal is to increase the user-friendliness of reproducibility tools for researchers.
While we have invested much effort to ensure this goal, the package is still in early development and undoubtedly has rough edges and even bugs that went undetected.
We hope that despite the early development stage, `repro` can support researchers to work more reproducible.

The workflow we present here introduces several software solutions which in itself presents a danger to reproducibility.
To profit from automatic and convenient reproduction researcher must install Git, Make, and Docker.
However, Git and Make are itself included in the Docker image created by `repro`.
Researchers can therefore employ the Docker image manually to download the Git repository and execute Make for full reproduction.
In other words the only requirement for reproduction and therefore Achilles' heel for this workflow is Docker.
The Docker approach has two vulnerabilities.
First, and more important, the Docker image for the project has to remain available.
The Dockerfile on the other hand is insufficient because it relies on to many service providers (e.g. Microsoft R Application Network, Ubuntu Package Archive).
The Docker image can be saved to disk and treated like any other file that requires careful archivation.
If the existence of the Docker image is guaranteed, researchers still require software to run the image.
To that end they can either rely on Docker itself or Docker-compatible alternatives (e.g. CoreOS rkt, Mesos Containerizer, Singularity).
The only way to remove the reliance on such external software is to turn the Docker image into a full operating system that subsequently can be installed and run on almost any modern computer.
This process is technically possible and would guarantee reproducibility for decades.
However, this process requires much technical knowledge and is currently not supported by `repro`.
In any case and without such precautions do the R Markdown, the Makefile and the Dockerfile provide information that allow researchers to manually trace the computational steps and recreate the computational environment.

@@

We realize that PACs are challenging.
Not every research question or project is suitable for preregistration in general, and many analytical strategies could prove unwieldy for a PAC.
Therefore, PAC should not be seen as an attempt to invalidate classical preregistration or not preregistered research but as a compliment.
Some may see the rigidness of PAC as to much of a good thing.
However, we are convinced that many reasons justify changing the preregistered code, but we are equally convinced that readers should judge themselves if they find these deviations from the preregistration meaningful.

Another challenge is that PAC requires more programming expertise than a usual analysis.
In contrast to a reproducible script, which should work on the same data, a PAC must work on different data.
We call the property of code that can run on different inputs from a similar context and produce valid outputs "reusability".
This property is based on reproducibility but requires the researcher to more carefully write the software [@lanerganSoftwareEngineeringReusable1989] such that it is _build-for-reuse_ [@al-badareenReusableSoftwareComponents2010].
The reproducible workflow we present here is heavily automated and hence promotes reusability.
Increased automation is more and more recognized as a means to improve the research process [@rouderMinimizingMistakesPsychological2019], and therefore this workflow fits well together with other innovations that employ automation, like machine-readable hypothesis tests [@lakensImprovingTransparencyFalsifiability2021] or automated data documentation [@arslanHowAutomaticallyDocument2019].
Automated and reusable research projects promise a wide range of applications, among them PAC [possibly submitted as a registered report @nosekRegisteredReports2014; @chambersWhatNextRegistered2019], Direct Replication[@simonsValueDirectReplication2014], fully automated living metanalysis [@elliottLivingSystematicReviews2014], Executable Research Articles [@elifesciencespublicationsELifeLaunchesExecutable2020], and other innovations like the live analysis of born open data [@rouderWhatWhyHow2016; @kekecsRaisingValueResearch2019].<!--AB very nice!!-->

@@

Please note that the analytical procedure in the PAC example of running different test procedures based on a fixed skewness threshold was created for didactic purposes, and it is not necessarily optimal for addressing your research question.
The t-test and Mann–Whitney–Wilcoxon test are arguably the most often used hypothesis test [@fagerlandTtestsNonparametricTests2012; @hortonStatisticalMethodsJournal2005 reports that 26% of all studies employ a t-test and 27% employ a rank based alternative in the New England Journal of Medicine in 2005].
The analytical strategy presented here is, in fact, suboptimal concerning several criteria (untested assumption of measurement invariance [@putnickMeasurementInvarianceConventions2016], underestimation of effect size in the presence of measurement error [@frostCorrectingRegressionDilution2000], overestimation of effect size for highly skewed distributions [@stonehouseRobustnessTestsCombined1998]).

@@

Open science practices are a continually evolving field where technical innovations foster changes in research practice.
Open data is much more widespread thanks to online storage facilities; preregistration are possible because there are preregistration platforms and so forth.
Similarly, we hope that fully automatic reproduction, e.g. with `repro` as technical innovation, increase scientific rigour and efficiency.
